Coherent Hyper-Raman Four Wave Mixing Vibrational Spectroscopy

### 2024-07-17 paper.tex
ed

\begin{abstract}
Nonlinear, four wave mixing vibrational spectroscopies are often used to probe intramolecular vibrational coupling and relaxation dynamics in isotropic media.
Three wave mixing vibrational spectroscopies are similarly used to understand the spectroscopy and dynamics of interfacial species.
Most of these methods rely on infrared or Raman transitions to generate output. 
Implementing a nonlinear spectroscopy involving hyper-Raman transitions, first appearing under third order perturbation theory, can provide a useful analogue to infrared and Raman based nonlinear spectroscopies to understand the spectroscopy and dynamics of isotropic systems.
Hyper Difference Frequency Generation (hyper-DFG, or HDFG) spectroscopy, also known as singly vibrationally enhanced spectroscopy (SIVE), is an underdeveloped, hyper-Raman based four wave mixing vibrational spectroscopy. 
Despite several experimental reports on HDFG, its spectroscopic properties have not been fully explored.
To this end, we explore the selection rules and response functions of HDFG spectroscopy.
Through the hyper-Raman $A,B,C$ terms, we explore the selection rules of singly and doubly resonant HDFG spectroscopy, and show HDFG is a hybrid IR-hyper-Raman technique.
HDFG is uniquely sensitive to non-Condon effects in molecular spectra.
Through a simple comparison test, HDFG output is shown to be similar in magnitude to vibrational sum frequency generation (vSFG).
Since HDFG can also be implemented as a two beam experiment, similar to vSFG, this makes HDFG a spectroscopy feasible for practitioners of vSFG.
Through a simple treatment of orientational averaging, HDFG provides a simple method to extract hyper-Raman hyperpolarizabilities ($\beta$).
HDFG shows promise as a method to disentangle vibrational spectra and dynamics in isotropic systems without need for anharmonicities.

@@

\section{Introduction}
Coherent multidimensional spectroscopy (CMDS) is a family of three and four wave mixing methods which form the optical analogue of multidimensional nuclear magnetic resonance (NMR) spectroscopy.\cite{RN103, Cho2008}
Multiresonant, four wave mixing CMDS experiments, first proposed by Oudar and Shen in 1980,\cite{RN307} directly probe coupling between different vibrational, electronic, and vibronic states. \cite{RN307, RN281, RN103, RN342, Cho2008, Ogilvie2019, RN325} 
CMDS has resolved anharmonicities, ultrafast dynamics, and other intramolecular couplings in numerous systems. \cite{RN342, RN343, RN324, Czech2015, Gaynor2017, Ogilvie2019, RN325}

Not all CMDS methods are used to dissect intramolecular couplings.\cite{Shen1987_CPL}
The most well known example of such a spectroscopy is vibrational sum frequency generation (vSFG), a three wave mixing technique used to probe vibrational structure and dynamics at interfaces. \cite{Shen1987_CPL, RN224}
Non-parametric, hyper difference frequency generation (hyper-DFG, or HDFG) spectroscopy, similar to difference frequency generation (\autoref{fig:comparisonwmel}), is a four wave mixing method which, when detuned from  electronic resonances, does not probe intramolecular couplings. 
HDFG was the first infrared CMDS method to successfully discriminate against non-resonant background.\cite{RN351, RN352}
HDFG has also been referred to as singly vibrationally enhanced (SIVE) spectroscopy. \cite{RN351}
In HDFG, an infrared pulse is resonant with a vibrational mode, and two other input pulses are used to stimulate a scattering process.
HDFG was documented long ago to have characteristics similar to both infrared spectroscopy and spontaneous hyper-Raman scattering. \cite{RN352}
Hyper-Raman scattering is the two photon analogue to Raman scattering. \cite{Cyvin1965, Terhune1965, Kozich2007}
Hyper-Raman scattering cross sections are several orders of magnitude smaller than Raman scattering cross sections, making it a difficult process to detect.\cite{RN515, Kelley2010} 
The different selection rules of hyper-Raman scattering, however, make it a unique alternative to Raman scattering to understand vibrational spectra and vibronic coupling in molecular systems.
Some years ago, Cho et al. proposed parametric six wave mixing techniques as the coherent analogue of hyper-Raman scattering. \cite{Cho1997, Cho1998}
However, since most six wave mixing methods cascade into four wave mixing processes and their homodyned output would be dependent upon the square magnitude of the hyper-Raman scattering cross section, \cite{RN243, RN515, Cho1997, Cho2000_Cascade} it is preferable to investigate coherent four wave mixing analogues to hyper-Raman scattering.

@@

While HDFG showed promise as an upconverted infrared spectroscopy method, it was quickly supplemented by a doubly vibrationally resonant method, doubly vibrationally enhanced (DOVE) spectroscopy, to investigate intra- and intermolecular vibrational coupling. \cite{RN345, RN101, Cho2000}
As a result, investigations of HDFG stagnated; to date, there are roughly a dozen studies which report on HDFG. \cite{Zilian1994, RN350, RN416, RN351, RN352, RN353, Chen1998, RN362, RN418, Bonn2024, McDonnell2024}
Recent reports highlight non-negligible HDFG interference in DOVE experiments as well as the suitability of HDFG pathways to interpret ultrafast relaxation dynamics in isotropic systems. \cite{Bonn2024, McDonnell2024}
These reports motivate deeper analysis on the properties of HDFG spectroscopy.

@@

\subsection{Placzek Method}
We first investigate the selection rules through a Placzek type formalism.
Contracting over the virtual states forms the hyper-Raman hyperpolarizability $\beta$.\cite{Long1970} 
Assuming $\omega_2$, $\omega_3$ are significantly detuned from any resonance (\autoref{fig:comparisonwmel}),\cite{Placzek1934, Long1970} \autoref{sivegamma} is written as 
\begin{equation}\label{sivebeta}
	\gamma_{ijkl}^{gv} =	-\frac{1}{\varepsilon_0} \frac{1}{4D \hbar}\frac{\beta^{gv}_{ijk} \mu^{gv}_{l}}{\Delta_{gv}} \rho_{gg}
\end{equation}
It is important to note that the symmetry properties of $\beta_{ijk}$ change dependent upon whether $\omega_2$ and $\omega_3$ are degenerate. \cite{Andrews1978, Altmann1982}
Taylor expanding the dipole and first hyperpolarizability operators in terms of an n$^{\text{th}}$ reduced normal mode coordinate $Q_n$ about equilibrium as\cite{Long1970, Shen90}
\begin{subequations}
	\begin{equation}
		\mu_l = \mu_{l,0} + \left(\frac{\partial \mu_l}{\partial Q_n}\right)_0 Q_n + \order{{Q_m Q_n}}
	\end{equation}
	\begin{equation}
		\beta_{ijk} = \beta_{ijk,0} + \left(\frac{\partial \beta_{ijk}}{\partial Q_n}\right)_0 Q_n + \order{{Q_m Q_n}}
	\end{equation}
\end{subequations}
and substituting into \autoref{sivebeta} gives the HDFG hyperpolarizability to $\order{Q_n}$ as \begin{equation}\label{SIVEselection}
	\gamma_{ijkl}^{gv} =	-\frac{1}{\varepsilon_0} \frac{1}{8D m_n \omega_{vg}}  \frac{1}{{\Delta_{gv}}} \ \left(\frac{\partial \beta^{gv}_{ijk}}{\partial Q_n}\right)_0 \left({\frac{\partial \mu^{gv}_{l}}{\partial Q_n}}\right)_0  \rho_{gg}
\end{equation}
where $\mel{v}{Q_n}{g} = \sqrt{\frac{\hbar}{2\omega_{vg}}}$.  \cite{RN230}
Since this expression is non-zero in the harmonic oscillator limit, HDFG output is allowed for harmonic transitions. 
Similar to SFG, HDFG does not demand anharmonic ground state potential energy surfaces for spectral output. \cite{Shen94, Cho2000}
This selection rule is generally valid for any HDFG or hyper sum frequency generation (HSFG) process in the $-\vec{k}_1 + \vec{k}_2  + \vec{k}_3$ and $\vec{k}_1 + \vec{k}_2  + \vec{k}_3$ phasematching geometries when $\omega_2$ and $\omega_3$ are sufficiently detuned from any resonance.

@@

By contracting over the virtual vibrational states $\ket{n}$, the $A$ and $B$ coefficients, where $B = B_1 + B_2$, are written as
\begin{widetext}
\begin{subequations}\label{ABterms}
\begin{equation}
	\begin{split}
		A_{ijk} = \frac{1}{\hbar^2}\sum_{m,e,v'} M^{ge}_{0,i} 
		M^{em}_{0,j} 
		M^{mg}_{0,k}
		 \langle v | v' \rangle
		 \langle v' | 0 \rangle 
		 \frac{1}{\Delta_{ev', g0}}
		 \\
	\end{split}
\end{equation}
	\begin{equation}
		\begin{split}
			B_{1_{ijk}} &= \frac{1}{\hbar^2} \sum_{m,e,v',z} M^{ge}_{0,i} \langle v | v' \rangle \left(
			 \frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k} \mel{v'}{Q_z}{0} 
			+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z} \mel{v'}{Q_z}{0} \right) \frac{1}{\Delta_{ev', g0}}\\
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
			B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,v',z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
			M^{mg}_{0,k} \mel{v}{Q_z}{v'} 
			\langle v' | 0 \rangle 
			\frac{1}{\Delta_{ev', g0}}
	\end{split}
	\end{equation}
\end{subequations}
\end{widetext}
where terms such as $\langle a | b \rangle$ and $\mel{a}{Q}{b}$ are Franck-Condon factors and Herzberg-Teller-type integrals, respectively. 
These expressions are valid for both singly and doubly resonant HDFG, but based upon the resonance schemes, further simplifications can be made (\textit{vide infra}).
\subsubsection{Singly Resonant HDFG}
In the case of singly resonant HDFG, $\omega_3$ is also detuned from any resonance, which permits contraction over ${v'}$ as the dependence on $\ket{v'}$ in the resonance denominator can be neglected. 
This limit is usually reached when $\abs{\omega_{ev',g0} - (\omega_2 + \omega_3)} \gg 2\Gamma_{ev',g0}$.
In such a limit, assuming a normal mode basis, we see for singly resonant HDFG that 
\begin{subequations}
\begin{equation}
	A_{ijk} \sim \sum_{v'}\langle v | v' \rangle \langle v' | 0 \rangle = \delta_{v0} = 0
\end{equation}
\begin{equation}
	B_{1_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,z} M^{ge}_{0,i} \left( 
	\frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k}
	+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z} \right) \frac{\mel{v}{Q_z}{0}}{\Delta_{e, g0}}
\end{equation}
\begin{equation}
	B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
	M^{mg}_{0,k}  
	\frac{\mel{v}{Q_z}{0}}{\Delta_{e, g0}}
\end{equation}
\end{subequations}
i.e., A$_{ijk}$ vanishes, but B$_{ijk}$ is non-zero, for singly resonant HDFG. 
When $\abs{\omega_{ev',g0} - (\omega_2 + \omega_3)} \gtrsim \Gamma_{ev',g0}$, $\Re(\frac{1}{\Delta_{ev', g0}})$ can allow for small $A$ term contributions.
As such, $A$ usually only vanishes for highly detuned electronic states. 
In the limit where $A$ vanishes, the singly resonant HDFG susceptibility can be approximated as $\chi^{(3)}_{IJKL} \sim \langle B_{ijk} \mu_l \rangle$, or, while a vibrational mode must be infrared active for HDFG output, vibronic activity is necessary for singly resonant HDFG response.
The breakdown of the Condon approximation in singly resonant HDFG supports the relatively small HDFG $\chi^{(3)}$ values ($\sim$ order of magnitude larger than nonresonant background) in hexane and similar organic solvents. \cite{RN350, RN351, RN353}

@@

The impact of vibronic coupling in the 2D spectrum is made clear by using a simple model.
Using a two-well system described by $H_g = p^2/2m + 1/2 \hbar \omega Q^2$ and $H_e = p^2/2m + 1/2 \hbar \omega (Q-\Delta)^2 +\hbar \omega_{eg}$, the A and B contributions can be easily evaluated using Franck-Condon and Herzberg-Teller integrals tabulated elsewhere. \cite{Carlson1988thesis} 
By working only in terms of a single normal mode (i.e. removing the sum over $z$) and assuming electronic transitions occur only between $|g), |e)$ and that $v =1$, $v' = \{0,1,2\}$, the A and B terms are rewritten as
\begin{widetext}
	\begin{subequations}\label{ABterms_DR}
		\begin{equation}
			\begin{split}
				A_{ijk} = \frac{1}{\hbar}\sum_{v'} M^{ge}_{0,i} 
				\Lambda^{eg}_{0,jk}
				\langle 1 | v' \rangle
				\langle v' | 0 \rangle 
				\frac{1}{\Delta_{ev',g0}}
				\\
			\end{split}
		\end{equation}
		\begin{equation}
			\begin{split}
				B_{1_{ijk}} &= \frac{1}{\hbar^2} \sum_{v',m} M^{ge}_{0,i} \langle 1 | v' \rangle \left(
				\frac{\partial M^{em}_{j}}{\partial Q} M^{mg}_{0,k} \mel{v'}{Q}{0} 
				+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q} \mel{v'}{Q}{0} \right) \frac{1}{\Delta_{ev', g0}}\\
			\end{split}
		\end{equation}
		\begin{equation}
			\begin{split}
				B_{2_{ijk}} = \frac{1}{\hbar} \sum_{v'} \frac{\partial M^{ge}_{i}}{\partial Q} 
				\Lambda^{eg}_{0,jk} 
				\mel{1}{Q}{v'} 
				\langle v' | 0 \rangle 
				\frac{1}{\Delta_{ev',g0}}
			\end{split}
		\end{equation}
	\end{subequations}
\end{widetext}
where the electronic two photon absorption operator, $\Lambda$ was invoked to simplify summation over virtual states. \cite{McClain1977}
It is clear from \autoref{ABterms_DR} that the Condon approximation breaks down for centrosymmetric molecules, as centrosymmetry invokes strict parity constraints that make states not accessible via both one and two photon transitions.
This makes DR-HDFG a unique tool for studying the electronic structure of centrosymmetric species in isotropic media, as it is sensitive to non-Condon effects. \cite{Olson2018}
A simulated spectrum highlighting the contributions of these terms (\autoref{fig:doubres_spec}) on potential wells with displacements $\Delta$ using arbitrary values listed in \autoref{T:hdfgvalues} show...
\begin{figure}[!htbp]
	\centering
	\includegraphics[width=3.375in]{drsive_spectrum.png}
	\caption{DR-HDFG $\gamma_{ijkl}$ spectrum for a simple two harmonic well system.
		(a) Potential energy surface, (b) 2D DR-HDFG spectrum for $\Delta = 0.5$.}
	\label{fig:doubres_spec}
\end{figure}

@@

Since the hyper-Raman process is stimulated by a two-photon process, a setup using only one optical parametric amplifier can perform a coherent four wave mixing hyper-Raman experiment by investigating $\pm \vec{k}_1 + 2\vec{k}_2$ processes (i.e. two laser CMDS), assuming a two photon absorption from a single pulse stimulates the hyper-Raman process.
The most commonly used two laser CMDS method is sum frequency generation, a $\chi^{(2)}$ technique whose phasematching is dictated by $\vec{k}_1 + \vec{k}_2$.
SFG output scales as $\chi^{(2)}_{IJK} \sim \langle \alpha_{ij} \mu_k \rangle$, where $\alpha_{ij}$ is the Raman polarizability tensor.
Since SFG is dependent upon transitions that transform like both rank-one and rank-two tensors, parity enforces a microscopic constraint that SFG cannot yield output for centrosymmetric species under the electric dipole approximation due to the presence of an inversion center. \cite{RN230}
Macroscopically, oddness in spatial inversion of the output polarization eliminates output from centrosymmetric species under the electric dipole approximation.\cite{RN133, RN132}
As a result, the output of SFG is significantly reduced relative to most third order spectroscopies because the output depends on surface species.
However, since HDFG is dependent upon a hyper-Raman transition, it is useful to compare the relative output of the two processes to identify how HDFG output compares to SFG.

>>>>

\begin{abstract}
Nonlinear, four wave mixing vibrational spectroscopies are often used to probe intramolecular vibrational coupling and relaxation dynamics in isotropic media.
Three wave mixing vibrational spectroscopies are similarly used to understand the spectroscopy and dynamics of interfacial species.
Most of these methods rely on infrared or Raman transitions to generate output. 
Implementing a nonlinear spectroscopy involving hyper-Raman transitions, first appearing under third order perturbation theory, can provide a useful analogue to infrared and Raman based nonlinear spectroscopies to understand the spectroscopy and dynamics of isotropic systems.
Hyper Difference Frequency Generation (hyper-DFG, or HDFG) spectroscopy, also known as singly vibrationally enhanced spectroscopy (SIVE), is an underdeveloped, hyper-Raman based four wave mixing vibrational spectroscopy. 
Despite several experimental reports on HDFG, its spectroscopic properties have not been fully explored.
To this end, we explore the selection rules and response functions of HDFG spectroscopy.
Through the hyper-Raman $A,B,C$ terms, we explore the selection rules of singly and doubly resonant HDFG spectroscopy, and show HDFG is a hybrid IR-hyper-Raman technique.
Since all infrared active modes have non-zero hyper-Raman activity, this makes HDFG a type of upconverted infrared spectroscopy.
HDFG is found to be uniquely sensitive to non-Condon effects in molecular spectra.
Through a simple comparison test, HDFG output is shown to be similar in magnitude to vibrational sum frequency generation (vSFG).
Since HDFG can also be implemented as a two beam experiment, similar to vSFG, this makes HDFG a spectroscopy feasible for practitioners of vSFG.
Through a simple treatment of orientational averaging, HDFG provides a simple method to extract hyper-Raman hyperpolarizabilities ($\beta$) of infrared active vibrations.
HDFG shows promise as a method to disentangle vibrational spectra and dynamics in isotropic systems without need for anharmonicities.

@@

\section{Introduction}
Coherent multidimensional spectroscopy (CMDS) is a family of three and four wave mixing methods which form the optical analogue of multidimensional nuclear magnetic resonance (NMR) spectroscopy.\cite{RN103, Cho2008}
Multiresonant, four wave mixing CMDS experiments, first proposed by Oudar and Shen in 1980,\cite{RN307} directly probe coupling and correlations between different vibrational, electronic, and vibronic states. \cite{RN307, RN281, RN103, RN342, Cho2008, Dong2015, Ogilvie2019, RN325} 
CMDS has resolved anharmonicities, ultrafast dynamics, and other intramolecular couplings in numerous systems. \cite{RN342, RN343, RN324, Czech2015, Gaynor2017, Ogilvie2019, RN325}

Not all CMDS methods are used to dissect intramolecular couplings.\cite{Shen1987_CPL}
The most well known example of such a spectroscopy is vibrational sum frequency generation (vSFG), a three wave mixing technique used to probe vibrational structure and dynamics at interfaces. \cite{Shen1987_CPL, RN224}
Non-parametric, hyper difference frequency generation (hyper-DFG, or HDFG) spectroscopy, similar to difference frequency generation (\autoref{fig:comparisonwmel}), is a four wave mixing method which, when detuned from  electronic resonances, does not probe intramolecular couplings. 
HDFG was the first infrared CMDS method to successfully discriminate against non-resonant background.\cite{RN351, RN352}
HDFG has also been referred to as singly vibrationally enhanced (SIVE) spectroscopy. \cite{RN351}
In HDFG, an infrared pulse is resonant with a vibrational mode, and two other input pulses are used to stimulate a scattering process.
HDFG was documented long ago to have characteristics similar to both infrared spectroscopy and spontaneous hyper-Raman scattering. \cite{RN352}
Hyper-Raman scattering is the two photon analogue to Raman scattering. \cite{Cyvin1965, Terhune1965, Kozich2007}
Hyper-Raman scattering cross sections are commonly several orders of magnitude smaller than Raman scattering cross sections, making hyper-Raman a difficult process to detect.\cite{RN515, Kelley2010} 
The different selection rules of hyper-Raman scattering make it a unique alternative to Raman scattering to understand vibrational spectra and vibronic coupling in molecular systems. \cite{Olson2018}
Some years ago, Cho et al. proposed parametric six wave mixing techniques as the coherent analogue of hyper-Raman scattering. \cite{Cho1997, Cho1998}
However, since most six wave mixing methods cascade into four wave mixing processes and their homodyned output would be dependent upon the square magnitude of the hyper-Raman scattering cross section, \cite{RN243, RN515, Cho1997, Cho2000_Cascade} it is preferable to investigate four wave mixing analogues to hyper-Raman scattering.

@@

While HDFG showed promise as an upconverted infrared spectroscopy method, it was quickly supplemented by a doubly vibrationally resonant method, doubly vibrationally enhanced (DOVE) spectroscopy, to investigate intra- and intermolecular vibrational coupling. \cite{RN345, RN101, Cho2000}
As a result, investigations of HDFG stagnated; to date, there are roughly a dozen studies which report on HDFG. \cite{Zilian1994, RN350, RN416, RN351, RN352, RN353, Chen1998, RN362, RN418, Bonn2024, McDonnell2024}
Recent reports highlight non-negligible HDFG interference in DOVE experiments as well as the suitability of HDFG pathways to interpret ultrafast relaxation dynamics in isotropic systems. \cite{Bonn2024, McDonnell2024}
These reports motivate deeper analysis of HDFG spectroscopy.

@@

\subsection{Placzek Method}
We first investigate the selection rules through a Placzek type formalism.
Contracting over the virtual states forms the hyper-Raman hyperpolarizability $\beta$.\cite{Long1970} 
Assuming $\omega_2$, $\omega_3$ are significantly detuned from any resonance (\autoref{fig:comparisonwmel}),\cite{Placzek1934, Long1970} \autoref{sivegamma} is written as 
\begin{equation}\label{sivebeta}
	\gamma_{ijkl}^{gv} =	-\frac{1}{\varepsilon_0} \frac{1}{4D \hbar}\frac{\beta^{gv}_{ijk} \mu^{gv}_{l}}{\Delta_{gv}} \rho_{gg}
\end{equation}
It is important to note that the symmetry properties of $\beta_{ijk}$ change dependent upon whether $\omega_2$ and $\omega_3$ are degenerate. \cite{Denisov1986}
Taylor expanding the dipole and first hyperpolarizability operators in terms of an n$^{\text{th}}$ reduced normal mode coordinate $Q_n$ about equilibrium as\cite{Long1970, Shen90}
\begin{subequations}
	\begin{equation}
		\mu_l = \mu_{l,0} + \left(\frac{\partial \mu_l}{\partial Q_n}\right)_0 Q_n + \order{{Q_m Q_n}}
	\end{equation}
	\begin{equation}
		\beta_{ijk} = \beta_{ijk,0} + \left(\frac{\partial \beta_{ijk}}{\partial Q_n}\right)_0 Q_n + \order{{Q_m Q_n}}
	\end{equation}
\end{subequations}
and substituting into \autoref{sivebeta} gives the HDFG hyperpolarizability to $\order{Q_n}$ as \begin{equation}\label{SIVEselection}
	\gamma_{ijkl}^{gv} =	-\frac{1}{\varepsilon_0} \frac{1}{8D m_n \omega_{vg}}  \frac{1}{{\Delta_{gv}}} \ \left(\frac{\partial \beta^{gv}_{ijk}}{\partial Q_n}\right)_0 \left({\frac{\partial \mu^{gv}_{l}}{\partial Q_n}}\right)_0  \rho_{gg}
\end{equation}
where $\mel{v}{Q_n}{g} = \sqrt{\frac{\hbar}{2\omega_{vg}}}$.  \cite{RN230}
Since this expression is non-zero in the harmonic oscillator limit, HDFG output is allowed for harmonic transitions. 
Similar to SFG, HDFG does not demand anharmonic ground state potential energy surfaces for spectral output. \cite{Shen94, Cho2000}
This selection rule is generally valid for any HDFG or hyper sum frequency generation (HSFG) process in the $-\vec{k}_1 + \vec{k}_2  + \vec{k}_3$ and $\vec{k}_1 + \vec{k}_2  + \vec{k}_3$ phasematching geometries when $\omega_2$ and $\omega_3$ are sufficiently detuned from any resonance.

@@

By contracting over the virtual vibrational states $\ket{n}$, the $A$ and $B$ coefficients, where $B = B_1 + B_2$, are written as
\begin{widetext}
\begin{subequations}\label{ABterms}
\begin{equation}
	\begin{split}
		A_{ijk} = \frac{1}{\hbar^2}\sum_{m,e,v'} M^{ge}_{0,i} 
		M^{em}_{0,j} 
		M^{mg}_{0,k}
		 \langle v | v' \rangle
		 \langle v' | 0 \rangle 
		 \frac{1}{\Delta_{ev', g0}}
		 \\
	\end{split}
\end{equation}
	\begin{equation}
		\begin{split}
			B_{1_{ijk}} &= \frac{1}{\hbar^2} \sum_{m,e,v',z} M^{ge}_{0,i} \langle v | v' \rangle \left(
			 \frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k} \mel{v'}{Q_z}{0} 
			+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z} \mel{v'}{Q_z}{0} \right) \frac{1}{\Delta_{ev', g0}}\\
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
			B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,v',z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
			M^{mg}_{0,k} \mel{v}{Q_z}{v'} 
			\langle v' | 0 \rangle 
			\frac{1}{\Delta_{ev', g0}}
	\end{split}
	\end{equation}
\end{subequations}
\end{widetext}
where terms such as $\langle a | b \rangle$ and $\mel{a}{Q}{b}$ are Franck-Condon factors and Herzberg-Teller-type integrals, respectively. 
These expressions are valid for both singly and doubly resonant HDFG, but based upon the resonance schemes, further simplifications can be made (\textit{vide infra}).
\subsubsection{Singly Resonant HDFG}
In the case of singly resonant HDFG, $\omega_3$ is also detuned from any resonance, which permits contraction over ${v'}$ as the dependence on $\ket{v'}$ in the resonance denominator can be neglected. 
This limit is usually reached when $\abs{\omega_{ev',g0} - (\omega_2 + \omega_3)} \gg 2\Gamma_{ev',g0}$.
In such a limit, assuming a normal mode basis, we see for singly resonant HDFG that 
\begin{subequations}
\begin{equation}
	A_{ijk} \sim \sum_{v'}\langle v | v' \rangle \langle v' | 0 \rangle = \delta_{v0} = 0
\end{equation}
\begin{equation}
	B_{1_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,z} M^{ge}_{0,i} \left( 
	\frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k}
	+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z} \right) \frac{\mel{v}{Q_z}{0}}{\Delta_{e, g0}}
\end{equation}
\begin{equation}
	B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
	M^{mg}_{0,k}  
	\frac{\mel{v}{Q_z}{0}}{\Delta_{e, g0}}
\end{equation}
\end{subequations}
i.e., A$_{ijk}$ vanishes, but B$_{ijk}$ is non-zero, for singly resonant HDFG. 
When $\abs{\omega_{ev',g0} - (\omega_2 + \omega_3)} \gtrsim \Gamma_{ev',g0}$, $\Re(\frac{1}{\Delta_{ev', g0}})$ can allow for small $A$ term contributions.
As such, $A$ usually only vanishes for highly detuned electronic states. 
In the limit where $A$ vanishes, the singly resonant HDFG susceptibility can be approximated as $\chi^{(3)}_{IJKL} \sim \langle B_{ijk} \mu_l \rangle$, or, while a vibrational mode must be infrared active for HDFG output, vibronic activity is necessary for singly resonant HDFG response.
The breakdown of the Condon approximation in singly resonant HDFG supports the relatively small HDFG $\chi^{(3)}$ values ($\sim$ order of magnitude larger than nonresonant background) for hexane and similar organic solvents. \cite{RN350, RN351, RN353}

@@

By working only in terms of a single normal mode (i.e. removing the sum over $z$) and assuming electronic transitions occur only between $|g), |e)$ and that $v =1$, $v' = \{0,1,2\}$, the A and B terms are rewritten as
\begin{widetext}
	\begin{subequations}\label{ABterms_DR}
		\begin{equation}
			\begin{split}
				A_{ijk} = \frac{1}{\hbar}\sum_{v'} M^{ge}_{0,i} 
				\Lambda^{eg}_{0,jk}
				\langle 1 | v' \rangle
				\langle v' | 0 \rangle 
				\frac{1}{\Delta_{ev',g0}}
				\\
			\end{split}
		\end{equation}
		\begin{equation}
			\begin{split}
				B_{1_{ijk}} &= \frac{1}{\hbar^2} \sum_{v',m} M^{ge}_{0,i} \langle 1 | v' \rangle \left(
				\frac{\partial M^{em}_{j}}{\partial Q} M^{mg}_{0,k} \mel{v'}{Q}{0} 
				+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q} \mel{v'}{Q}{0} \right) \frac{1}{\Delta_{ev', g0}}\\
			\end{split}
		\end{equation}
		\begin{equation}
			\begin{split}
				B_{2_{ijk}} = \frac{1}{\hbar} \sum_{v'} \frac{\partial M^{ge}_{i}}{\partial Q} 
				\Lambda^{eg}_{0,jk} 
				\mel{1}{Q}{v'} 
				\langle v' | 0 \rangle 
				\frac{1}{\Delta_{ev',g0}}
			\end{split}
		\end{equation}
	\end{subequations}
\end{widetext}
where the electronic two photon absorption operator, $\Lambda$ was invoked to simplify summation over virtual states. \cite{McClain1977}
It is clear from \autoref{ABterms_DR} that the Condon approximation breaks down for centrosymmetric molecules, as centrosymmetry invokes strict parity constraints that make states not accessible via both one and two photon transitions.
This makes DR-HDFG a unique tool for studying the electronic structure of centrosymmetric species in isotropic media, as it is sensitive to non-Condon effects. \cite{Olson2018}
The lineshapes of electronic-vibrational coupling probed by DR-HDFG can inform on electronic-vibrational correlations. \cite{Carlson1990line,Dong2015}
These effects are under investigation and will the subject of a future publication.

The impact of vibronic coupling in the 2D spectrum is made clear by using a simple model.
Using a two-well system described by $H_g = p^2/2m + 1/2 \hbar \omega Q^2$ and $H_e = p^2/2m + 1/2 \hbar \omega (Q-\Delta)^2 +\hbar \omega_{eg}$, the A and B contributions can be easily evaluated using Franck-Condon and Herzberg-Teller integrals tabulated elsewhere. \cite{Carlson1988thesis} 
A simulated spectrum highlighting the contributions of these terms (\autoref{fig:doubres_spec}) on potential wells with $\Delta = 0.5$ using values listed in \autoref{T:hdfgvalues} dissects the $A$ and $B$ contributions to the DR-SIVE spectrum.

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=3.375in]{drsive_spectrum.png}
	\caption{DR-HDFG $\gamma_{ijkl}$ spectrum for a simple two harmonic well system.
		(a) Potential energy surface, (b) 2D DR-HDFG spectrum for $\Delta = 0.5$.}
	\label{fig:doubres_spec}
\end{figure}

@@

Since the hyper-Raman process is stimulated by a two-photon process, a setup using only one optical parametric amplifier can perform a coherent four wave mixing hyper-Raman experiment by investigating $-\vec{k}_1 + 2\vec{k}_2$ processes (i.e. two laser CMDS), assuming a two photon absorption from a single pulse stimulates the hyper-Raman process.
The most commonly used two laser CMDS method is vibrational sum frequency generation (vSFG), a $\chi^{(2)}$ technique whose phasematching is dictated by $\vec{k}_1 + \vec{k}_2$.
SFG output scales as $\chi^{(2)}_{IJK} \sim \langle \alpha_{ij} \mu_k \rangle$, where $\alpha_{ij}$ is the Raman polarizability tensor.
Both processes can be phasematched in media with normal indicies of refraction, unlike the $\vec{k}_1 + 2\vec{k}_2$ process.\cite{RN120}
Macroscopically, oddness in spatial inversion of the vSFG polarization eliminates output from centrosymmetric species under the electric dipole approximation.\cite{RN132, RN133}
As a result, the output of vSFG is significantly reduced relative to most third order spectroscopies because the output depends on surface number density, many orders of magnitude smaller than the bulk surface density. 
However, since HDFG is dependent upon a hyper-Raman transition, which is several orders of magnitude weaker than Raman transitions,\cite{RN515} it is useful to identify how HDFG output compares to vSFG.
